---
title: "HW02 - Basics of Data Frames"
output: github_document
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```


**2) Importing the data 2 ways**

Base
```{r}
dat1 <- read.csv(file = "C:/Users/13noa/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", colClasses = c("character", "character", NA, "character", "double", NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA))

str(dat1)

```

readr
```{r}
dat2 <- read_csv("C:/Users/13noa/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", col_types = cols(Position = col_factor(levels = c("C", "PF", "PG", "SF", "SG")), Salary = col_double()))

str(dat2)
```

```{r}
dat1$Experience <- replace(dat1$Experience, dat1$Experience == "R", "0")

dat1$Experience <- as.integer(dat1$Experience)

```

```{r}
dat1 <- mutate(dat1, Missed_FG = FGA-FGM, Missed_FT = FTA-FTM, PTS = FTM + 2 * Points2 + 3 * Points3, 
       REB = OREB + DREB, MPG = MIN/GP, EFF = (PTS + REB + AST + STL + BLK - Missed_FG - Missed_FT - TO)/GP)

summary(dat1$EFF)
hist(dat1$EFF)
```

```{r}
Top10 <- 
  slice(
    arrange(
      select(dat1, Player, Team, Salary, EFF),
            desc(EFF)),
        1:10)

Top10
```

```{r}
negativeEFF <- filter(dat1, EFF < 0)

negativeEFF
```

```{r}

EFF.cor <- c(cor(dat1$PTS, dat1$EFF), cor(dat1$REB, dat1$EFF), cor(dat1$AST, dat1$EFF), cor(dat1$STL, dat1$EFF), cor(dat1$BLK, dat1$EFF), cor(-dat1$Missed_FT, dat1$EFF), cor(-dat1$Missed_FG, dat1$EFF), cor(-dat1$TO, dat1$EFF))

```
```{r}
barplot(c(cor(dat1$PTS, dat1$EFF),
cor(dat1$REB, dat1$EFF),
cor(dat1$AST, dat1$EFF),
cor(dat1$STL, dat1$EFF),
cor(dat1$BLK, dat1$EFF),
cor(-dat1$Missed_FT, dat1$EFF),
cor(-dat1$Missed_FG, dat1$EFF),
cor(-dat1$TO, dat1$EFF)), names.arg = c("PTS", "REB", "AST", "STL", "BLK", "Missed_FT", "Missed_FG", "TO"), ylim = c(-1.0, 1.0), cex.names = 0.70, col = c("grey", "grey", "grey", "grey", "grey", "orange", "orange", "orange"), main = "Correlations between Player Stats and EFF")
```

```{r}
ggplot(dat1, aes(x = EFF, y = Salary)) + geom_point() + geom_smooth()
```
<br>Correlation Coefficient r = `r cor(dat1$Salary, dat1$EFF)`

There appears to be a somewhat strong positive correlation between the efficiency statistic and a player's salary.

```{r}
players2 <- filter(dat1, MPG >= 20)
ggplot(players2, aes(x = EFF, y = Salary)) + geom_point() + geom_smooth()
```
<br>Correlation Coefficient r = `r cor(players2$Salary, players2$EFF)`

Surprisingly, the linear correlation between a player's salary and their respective efficiency statistic is weaker when only including players with 20 or more MPG. I suppose this could be attributed to fatigue, i.e. the more minutes a player plays, the less efficient they become due to fatigue. Either that, or I made a mistake somewhere along the way.

##Reflections

Overall I'd say this assignment was simple, yet time consuming for me. I found myself having difficulties with little parts of certain functions. For example, I spent over an hour just trying to figure out how to read the data from the csv in as a factor for the *position* column. I also found some inelegant solutions to some things (writing NA a bunch of times while using the read.csv function) and I definitely want to work on making my code more tidy. 

