---
title: "hw03-noah-weingarten"
output: github_document
---


##**Basic Rankings**

```{r}
#Packages
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
#loading the data
teams <- read_csv(file = "../data/nba2017-teams.csv")
```


####**Ranking teams according to salary**

```{r}
ggplot(teams, aes(x = reorder(team, salary), y = salary)) +
  geom_bar(position = "dodge", stat = "identity", alpha = 0.6) +
  coord_flip() +
  scale_y_continuous(name = "Salary (in millions)", breaks = c(0, 40, 80, 120), limits = c(0, 130)) +
  geom_hline(yintercept = mean(teams$salary), size = 2, color = "red", alpha = 0.5) +
  ggtitle("NBA Teams ranked by Total Salary") +
  xlab("Team") +
  theme(plot.title = element_text(hjust = 0.5))
```

####**Ranking teams according to points scored**

```{r}
ggplot(teams, aes(x = reorder(team, points), y = points)) +
  geom_bar(position = "dodge", stat = "identity", alpha = 0.6) +
  coord_flip() +
  scale_y_continuous(name = "Total Points", breaks = c(0, 2500, 5000, 7500), limits = c(0, 10000)) +
  geom_hline(yintercept = mean(teams$points), size = 2, color = "red", alpha = 0.5) +
  ggtitle("NBA Teams ranked by Total Points") +
  xlab("Team") +
  theme(plot.title = element_text(hjust = 0.5))
```


####**Ranking teams according to total team efficiency**

```{r}
ggplot(teams, aes(x = reorder(team, efficiency), y = efficiency)) +
  geom_bar(position = "dodge", stat = "identity", alpha = 0.6) +
  coord_flip() +
  scale_y_continuous(name = "Total Team Efficiency", breaks = c(0, 50, 100, 150), limits = c(0, 200)) +
  geom_hline(yintercept = mean(teams$efficiency), size = 2, color = "red", alpha = 0.5) +
  ggtitle("NBA Teams ranked by Total Efficiency") +
  xlab("Team") +
  theme(plot.title = element_text(hjust = 0.5))
```

####**Description of the obtained rankings so far**

>For the most part, the teams with the higher salaries score more points
>and are more efficient, but there are some clear exceptions to this.
>One team that seems to get the most out of the money they spend is the 
>Philadelphia 76ers. They have the lowest total team salary and rank near
>the bottom in total points scored, but their total team efficiency is 
>among the best in the entire NBA. Without knowing much about the sport 
>of basketball, this leads me to believe that the 76ers are very efficient
>at things like rebounding, creating turnovers, and overall playing good defense.
>Despite how efficient they may be on defense, you can't win games if you don't score
>which is supported by the fact that they had the second worse record in their
>conference. This shows that while efficiency is a useful statistic that attempts
>to provide a clear picture of a teams ability, it is far from perfect.

##**Principal Components Analysis (PCA)**

```{r}
#Eigenvalues

teams.pca <- teams[c("points3", "points2", "free_throws", "off_rebounds", "def_rebounds",
                     "assists", "steals", "blocks", "turnovers", "fouls")]

pca <- prcomp(teams.pca, scale. = TRUE)

eigs <- data.frame(
  eigenvalue = pca$sdev^2,
  proportion = round((pca$sdev^2) / sum(pca$sdev^2), 4)
)

eigs$cumprop <- cumsum(eigs$proportion)
eigs

```

####**PCA Plot for PC1 and PC2**

```{r}
PC1 <- pca$x[ , 1]
PC2 <- pca$x[ , 2]

ggplot(data = teams, aes(x = PC1, y = PC2)) + geom_text(label = teams$team) +
  labs(x = "PC1", y = "PC2", title = "PCA plot (PC1 and PC2)") +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0)

```

####**Index based on PC1**

```{r}
PC1_transformed <- 100 * (PC1 - min(PC1))/(max(PC1) - min(PC1))

```

####**Ranking teams according to scaled PC1**

```{r}
ggplot(teams, aes(x = reorder(team, PC1_transformed), y = PC1_transformed)) +
  geom_bar(position = "dodge", stat = "identity", alpha = 0.6) +
  coord_flip() +
  scale_y_continuous(name = "First PC (Scaled from 1 to 100)", breaks = c(0, 25, 50, 75), limits = c(0, 100)) +
  ggtitle("NBA Teams ranked by Scaled PC1") +
  xlab("Team") +
  theme(plot.title = element_text(hjust = 0.5))
```

>Ranking the teams with the PC1 Index tells a slightly different story than the previous rankings.
>The biggest thing that jumps out to me is the ranking of the Cleveland Cavaliers.
>The Eastern Conference Champions are the 6th lowest team as far as PC1 can tell.
>Aside from that, the Warriors are far and away the number 1 ranked team,
>and the bottom 3 teams are separated from the rest to a more extreme degree.
>These things may be due to the way the data is scaled.

##**Comments and Reflections**

>This was the first time working on a project with this structure for me.
>I am definitely grateful for the experience.
>While it was difficult and time consuming at stages, I feel like I'm learning something
>very applicable with this class.

>Using relative paths makes everything much easier for the most part.

>This was my first time using an R script as well. I normally just use R Markdown.
>I can definitely see the benefit to using scripts because it makes the final product cleaner.

>I wouldn't say that anything in particular was hard or easy.
>The difficulty was right at what I expect it to be.

>I didn't receive any help in completing the assignment.

>I'd say it took me about 10 hours over the course of 2 days to complete.

>The most time consuming part was probably creating the first graph because
>I was unfamiliar with ggplot2.
>But after the general structure of the first graph was set, the other ones went much quicker.

>I love sports, and while I'm nothing more than a casual fan of basketball,
>advanced metrics have been an interest of mine for a while now.
>I'm definitely grateful for the chance to learn how to actually work with some.

