---
title: "Delving Deeper into 'dplyr'"
output: github_document
author: Noah Weingarten
---


###Introduction

Data wrangling is often one of the very first things one learns when beginning to code using R; and 'dplyr' is a package that is often used for this task. On the surface, the functions provided through this package are useful tools to have access to. You think, *"Of course this package will come in handy whenever I need to manipulate a data set!"* Indeed, this is similar to my thought process when I was first introduced to 'dplyr'. But as I've progressed through the earlier stages of this class, I've come to recognize that 'dplyr' and the data wrangling functions therein are arguably the most vital part of an R-programmer's repertoire. My hope is that, after reading this post, you feel inspired to re-visit the basics of data wrangling and gain mastery of the 'dplyr' package.

>"As the variety of data continues to increase, data wrangling will become more important and having good tools to do this will be a key differentiator for data scientists."<center> *Amyn Rajan, GM of Connectivity Solutions for Magnitude Software, inc.*


###Data Example

To illustrate just how important 'dplyr' is for an R-programmer, let's take a look at a data set that I have personally used and wrangled in the past. A friend of mine runs a catering service that caters lunches for multiple schools in the East Bay and this is an example of the raw data that she may receive through the website where parents order the food for their kids.

```{r, echo = FALSE, message = FALSE}
library(readr)
library(dplyr)

data <- read_csv("C://Users/13noa/Downloads/distribution-list-all.csv", skip = 5)
head(data.frame(data), 10)

```

Pretty gnarly, right? I mean, fruit is considered an *addon* to your child's order of *please choose a side*? It's convoluted and nonsensical to say the least, and that's just one example. This is what raw data is often like, and without competent data wrangling, it can be flat-out unusable.

###The basics of dplyr

dplyr breaks everything down into 5 primary functions that can be used to solve the majority of your data wrangling problems.

`select()` selects only the columns you specify<br>
`filter()` filters the data to include cases that fit the criteria you specify<br>
`arrange()` changes the order of the rows<br>
`mutate()` adds new variables that are functions of existing variables<br>
`summarise()` summarises the values to a single value

Finally, the almighty pipe operator `%>%` provides an intuitive way of chaining multiple dplyr commands together.

These 5 functions are very simple and straightforward and I will now use them to clean up our data.

```{r}
#Many of the variables included in the raw data are irrelevant.
#We can use select() to select the useful ones

data.1 <- data %>%
  select(c(2, 3, 7, 9))

head(data.frame(data.1), 10)

#Now that we are working with fewer variables, let's use filter()
#to select only the cases where the parent orders a side for their child.

data.2 <- data.1 %>%
  filter(Item == "Please Choose a Side")
head(data.frame(data.2), 10)

#So we've narrowed the data to only include side orders.
#We will use arrange() to order the Addons column alphabetically

data.3 <- data.2 %>%
  arrange(Addons)

head(data.frame(data.3), 10)

#With this we have all the different side orders grouped together.
#The names are quite confusing. Let's use mutate() to create a new
#variable called 'Side' to differentiate it from the other addons.
#Then we can use select() again to get rid of the Addons and Item 
#columns as they are no longer needed

data.4 <- data.3 %>%
  mutate("Side" = Addons) %>%
  select(c(1, 2, 5))

head(data.frame(data.4), 10)

#Finally, let's use summarise() to display the sides with more
#reasonable names.

```


