---
title: "Homework 1"
output: github_document
---

```{r}
load("nba2017-salary-points.RData")
new.salary <- salary/1000000
new.experience <- replace(experience, experience == "R", "0")
new.position <- factor(x = position, c("C", "PF", "SG", "PG", "SF"), c("Center", "Power Forward", "Shooting Guard", "Point Guard", "Small Forward"))

table(new.position)
```
```{r}
plot(points, new.salary)
```

There does seem to be a positive trend in the data. The majority of the points are clustered around low salary and low point totals.

```{r}
n <- 441
xbar <- sum(points)/n
ybar <- sum(new.salary)/n
varx <- (1/(n-1))*sum((points-xbar)^2)
vary <- (1/(n-1))*sum((new.salary-ybar)^2)
sdx <- sqrt(varx)
sdy <- sqrt(vary)
covxy <- (1/(n-1))*sum((points-xbar)*(new.salary-ybar))
corrxy <- covxy/(sdx*sdy)
n
xbar
ybar
varx
vary
sdx
sdy
covxy
corrxy

```

```{r}
b1 <- corrxy*(sdy/sdx)
b0 <- ybar-b1*xbar
yhat <- b0+b1*points

summary(yhat)

```

The regression equation is yhat = `r b0` + `r b1`x

The slope positive which indicates that salaries increase as points increase. Literally, I think it means that a player is predicted to earn approximately $`r round(1000000*b1, digits = 2)` per point.

The intercept is the amount a player in the NBA is predicted to earn should they score 0 points.
_____________________________________________________________________________________________

**Note: this happens when I try to use inline code to answer this part of Q4. Not sure how to fix

The predicted salary for a player who scores...

...0 points is $`r round(1000000*b0, digits = 2)`
_____________________________________________________________________________________________

```{r}
y0 <- (b0+b1*0)*1000000
y100 <- (b0+b1*100)*1000000
y500 <- (b0+b1*500)*1000000
y1000 <- (b0+b1*1000)*1000000
y2000 <- (b0+b1*2000)*1000000
```

The predicted salary for a player who scores...

...0 points is $`r y0`.

...100 points is $`r y100`.

...500 points is $`r y500`.

...1000 points is $`r y1000`.

...2000 points is $`r y2000`.


```{r}
plot(points, new.salary)
abline(b0, b1, col = "blue", lwd = 2)
text(2200, 18, labels = "regression", col = "blue")
text(2200, 29, labels = "lowess", col = "red")
lines(lowess(points, new.salary), col = "red", lwd = 2)
```

```{r}
#Residuals
ei <- new.salary - yhat
RSS <- sum(ei^2)
TSS <- sum((new.salary - ybar)^2)

rsq <- 1 - (RSS/TSS)
rsq

```

```{r}
#Scatterplot for experience to salary
plot(new.experience, new.salary)
lines(lowess(new.experience, new.salary), col = "red", lwd = 2)
```

```{r}
#3d scatterplot
library(scatterplot3d)

scatterplot3d(points, new.experience, new.salary, color = "red", main = "3D Scatterplot")
```

```{r}
#boxplot
boxplot(new.salary ~ position)
```

The "experience to salary" scatterplot shows a positive relationship between the two indicating that a more experienced player will earn a higher salary. 

The 3d scatterplot is difficult for me to interpret but it seems to support what the previous scatterplots have told us. Both experience and points scored are factors in determining a player's salary.

Based on the box plots, I would argue that position doesn't have much of an effect on salary considering the median value for each position is virtually equal. 
